---
title: "StatisticalInferenceProjectPart2"
author: "Ronda Hutchin"
date: "8/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part II Assignment Instructions
Analyze the ToothGrowth data in the R datasets package.
(1) Load the ToothGrowth data and perform some basic exploratory data analyses
(2) Provide a basic summary of the data.
(3) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
(4) State your conclusions and the assumptions needed for your conclusions.

## Step (1) Examine the data and using STR, HEAD and SUMMARY. To save space the results are hidden.
```{r results='hide'}
library(datasets)
library(ggplot2)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```


## Step (2) Basic Summary of data using a chart
Using ggplot, the Tooth Length versus Dose is charted. The tooth length increases as the dose increases.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
MyGrowthPlot <- ggplot(ToothGrowth, aes(x=dose, y=len,fill=dose)) +  geom_boxplot() +  ggtitle("Tooth Growth by Dose") + xlab("Dose in mg") + ylab("Tooth Length")
MyGrowthPlot

```
Using ggplot, the Tooth Length versus Delivery Method is higher for Orange Juice than for Vitamin C.  
```{r}
MyMethodPlot <- ggplot(ToothGrowth, aes(x=supp, y=len,fill=supp)) +  geom_boxplot() +  ggtitle("Tooth Growth by Delivery Method") + xlab("Delivery Method") + ylab("Tooth Length")
MyMethodPlot
```
## Step (3) Use Confidence Intervals and/or Hypothesis tests to compare Tooth Growth by Supp and Dose.


